---
title: "Group_Project"
author: "Yueyang Zhang"
date: "2019Äê12ÔÂ4ÈÕ"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
# Loading packages needed in following steps
library("tidyverse")
library(haven)
library(dplyr)
library(tidyr)
library(ResourceSelection)
library(ggplot2)
library(foreign)#
library(nnet)#
library(ggplot2)
library(reshape2)
library(lmerTest)
library(car)
library(nlme)

multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  ## A function used to plot several plots on the same page.
  ## found this func from internet
  ## input: ggplot item
  ## output: just plot
  
  require(grid)
  
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  
  numPlots = length(plots)
  
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  
  if (numPlots==1) {
    print(plots[[1]])
    
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

Here need to states how we deal with our data.(important)
```{r}
library(MASS)
detach("package:MASS", unload=TRUE)
# Load data and select variables we need and drop NA
X<-read_xpt("https://wwwn.cdc.gov/Nchs/Nhanes/2005-2006/DEMO_D.XPT")
X_variable<-X%>%select("SEQN","RIAGENDR","RIDAGEYR","DMDEDUC2","RIDRETH1")%>%
  drop_na()%>%
  filter(RIDAGEYR>=20,DMDEDUC2!=7,DMDEDUC2!=9)%>%
  mutate(RIAGENDR=as.numeric(RIAGENDR==1))%>%
  transmute(SEQN,gender=RIAGENDR,age=RIDAGEYR,race=RIDRETH1,education=DMDEDUC2)

health_insurance<-read_xpt("https://wwwn.cdc.gov/Nchs/Nhanes/2005-2006/HIQ_D.XPT")
health_insurance<-health_insurance%>%select(SEQN,HIQ011)%>%
  drop_na()%>%
  filter(HIQ011!=7,HIQ011!=9)%>%
  mutate(insurance=as.numeric(HIQ011==1))%>%
  select(SEQN,insurance)

Smoking<-read_xpt("https://wwwn.cdc.gov/Nchs/Nhanes/2005-2006/SMQ_D.XPT")
Smoking<-Smoking%>%
  select(SEQN,SMQ020)%>%
  drop_na()%>%
  filter(SMQ020<7)%>%
  mutate(smoking=as.numeric(SMQ020!=1))%>%
  select(SEQN,smoking)

BMI<-read_xpt("https://wwwn.cdc.gov/Nchs/Nhanes/2005-2006/BMX_D.XPT")
BMI<-BMI%>%select(SEQN,BMXBMI)%>%
  drop_na()%>%
  mutate(BMI=as.numeric(BMXBMI>=18.5&BMXBMI<=24.9))%>%
  select(SEQN,BMI)

Blood_pressure<-read_xpt("https://wwwn.cdc.gov/Nchs/Nhanes/2005-2006/BPX_D.XPT")
Blood_pressure<-Blood_pressure%>%select(SEQN,BPXSY1,BPXSY2,BPXSY3,BPXDI1,BPXDI2,BPXDI3)%>%
  gather(condition, BPX, BPXSY1:BPXDI3)%>%
  mutate(condition=substring(condition,1,5))%>%
  group_by(SEQN,condition)%>%
  summarise(BPX=mean(BPX,na.rm=T))%>%
  ungroup()%>%
  spread(condition,BPX)%>%
  drop_na()%>%
  filter(BPXDI!=0,BPXSY!=0)%>%
  transmute(SEQN,Blood_pressure=as.numeric((BPXDI<80)&(BPXSY<120)))

Diet_raw<-read_xpt("https://wwwn.cdc.gov/Nchs/Nhanes/2005-2006/DBQ_D.XPT")
Diet<-Diet_raw%>%select(SEQN,DBQ700)%>%
  drop_na()%>%
  filter(DBQ700!=7,DBQ700!=9)%>%
  transmute(SEQN,Diet=as.numeric(DBQ700<=3))

Diet_alt<-Diet_raw%>%
  select(SEQN,DBQ780)%>%
  drop_na()%>%
  filter(DBQ780!=77,DBQ780!=99)%>%
  transmute(SEQN,Diet=as.numeric(DBQ780<=4))

Physical_Activity<-read_xpt("https://wwwn.cdc.gov/Nchs/Nhanes/2005-2006/PAQIAF_D.XPT")
Physical_Activity<-Physical_Activity%>%
  select(SEQN,PADLEVEL,PADTIMES,PADDURAT)%>%
  drop_na()%>%
  mutate(times=PADTIMES*PADDURAT*PADLEVEL)%>%
  group_by(SEQN)%>%
  summarise(phy_act=as.numeric(sum(times)>=600))%>%
  select(SEQN,phy_act)

Blood_Cholesterol<-read_xpt("https://wwwn.cdc.gov/Nchs/Nhanes/2005-2006/TCHOL_D.XPT")
Blood_Cholesterol<-Blood_Cholesterol%>%
  select(SEQN,LBXTC)%>%
  drop_na()%>%
  transmute(SEQN,blood_cho=as.numeric(LBXTC<200))

Blood_Glucose<-read_xpt("https://wwwn.cdc.gov/Nchs/Nhanes/2005-2006/GLU_D.XPT")
Blood_Glucose<-Blood_Glucose%>%
  select(SEQN,LBXGLU)%>%
  drop_na()%>%
  transmute(SEQN,blood_glu=as.numeric(LBXGLU<=100))

# merge all seperate datasets together by SEQN
raw_data<-X_variable%>%inner_join(health_insurance, by = "SEQN")%>%
  inner_join(Smoking, by = "SEQN")%>%
  inner_join(BMI, by = "SEQN")%>%
  inner_join(Blood_pressure, by = "SEQN")%>%
  inner_join(Diet, by = "SEQN")%>%
  inner_join(Physical_Activity, by = "SEQN")%>%
  inner_join(Blood_Cholesterol, by = "SEQN")%>%
  inner_join(Blood_Glucose, by = "SEQN")

data<-raw_data%>%transmute(SEQN,CVH=smoking+Blood_pressure+phy_act+blood_cho+blood_glu+BMI+Diet,smoking,Blood_pressure,phy_act,blood_cho,blood_glu,BMI,Diet,gender,age,race,education,insurance)

# Then we get our final version dataset
data
```


```{r}
# First we analyze the relationship between gender and each facor of CVH score using logistic model
gender_smoking <- summary(glm(smoking~gender+education+age+insurance+race,data=data, family = "binomial"))
gender_BP <-summary(glm(Blood_pressure~gender+education+age+insurance+race,data, family = "binomial"))
gender_phy <- summary(glm(phy_act~gender+education+age+insurance+race,data, family = "binomial"))
gender_BC <- summary(glm(blood_cho~gender+education+age+insurance+race,data, family = "binomial"))
gender_BG <- summary(glm(blood_glu~gender+education+age+insurance+race,data, family = "binomial"))
gender_BMI <- summary(glm(BMI~gender+education+age+insurance+race,data, family = "binomial"))
gender_Diet <- summary(glm(Diet~gender+education+age+insurance+race,data, family = "binomial"))

seperate<-data.frame(factor=c("smoking","Blood_pressure","phy_act","blood_cho","blood_glu","BMI","Diet"),gender_effect=rep(0,7),p_value=rep(0,7),significance=rep("*",7),stringsAsFactors = FALSE)
j=1
for (i in list(gender_smoking,gender_BP,gender_phy,gender_BC,gender_BG,gender_BMI,gender_Diet)){
  seperate$gender_effect[j]=i$coefficients[2,1]
  seperate$p_value[j]=i$coefficients[2,4]
  p=rank(c(i$coefficients[2,4],0.001,0.01,0.05,0.1))[1]
  seperate$significance[j]=switch(p,
                                  "***",
                                  "**",
                                  "*",
                                  ".",
                                  " ")
  j=j+1
}

formattable::formattable(seperate)
```

Then we will conduct OLS analysis
```{r warning=FALSE}
# We begin first with OLS regression and some diagnostics to view the general relationship between our data.
OLS_full<-lm(CVH~gender+race+education+insurance+age,data)
summary(OLS_full)

# we delete race variable and get a seemly good model.
OLS_opt<-lm(CVH~gender+education+insurance+age,data)
summary(OLS_opt)

OLS2<-lm(CVH+1~gender+education+insurance+age,data)
library(MASS)
boxcox(OLS2,plotit=T)# 1 is in the confidence interval so no need to do transformation

dat=data.frame(fitted.values=as.vector(OLS_opt$fitted),residuals=as.vector(OLS_opt$residuals))
ggplot(data=dat,aes(x=fitted.values,y=residuals))+geom_point(color="red",alpha=0.1)+geom_smooth(se=T)
# It is obvious that OLS model doesn't fit well with our dependent variables discontinuous. But we can still obtain the information that among the 5 predictors, gender, education level, insurance status, and age are more significant that race.

```


```{r}
# We polt boxplots for response variable CVH grouped by different predictors. (Here we group age variable)
data$age_group=ceiling(data$age/10)-3
data$age_group[data$age==-1]=0

p1<-qplot(factor(gender), CVH, facets = . ~ factor(gender), 
      colour = factor(gender), geom = "boxplot", data = data)
p2<-qplot(factor(insurance), CVH, facets = . ~ factor(insurance), 
      colour = factor(insurance), geom = "boxplot", data = data)
p3<-qplot(factor(education), CVH, facets = . ~ factor(education), 
      colour = factor(education), geom = "boxplot", data = data)
p4<-qplot(factor(age_group), CVH, facets = . ~ factor(age_group), 
      colour = factor(age_group), geom = "boxplot", data = data)

multiplot(p1, p2, p3, p4,cols=2)

```
From the plots We can see that the CVH shows difference in different groups. It is resasonable to establish the following mixed effect model
```{r}
mixed=lme(CVH~gender+insurance+age+education, random=~1|age_group/education/insurance,  
         method = 'ML', data = data)

# Conduct Analysis of Variance and find this model dignificant.(?) and draw residuals_fitted plot
Anova(mixed)

dat=data.frame(fitted.values=as.vector(fitted(mixed)),residuals=as.vector(residuals(mixed)))
ggplot(data=dat,aes(x=fitted.values,y=residuals))+geom_point(color="red",alpha=0.1)+geom_smooth(se=T)

```
Next we will test whether random effects are warranted
```{r}
# lm.test
dev1 = -2*logLik(mixed);dev0 = -2*logLik(OLS_opt)
devdiff = as.numeric(dev0-dev1)
dfdiff <- attr(dev1,"df")-attr(dev0,"df"); 
cat('Chi-square =', devdiff, '(df=', dfdiff,'), p =', 
    pchisq(devdiff,dfdiff,lower.tail=FALSE))
```

And we also test the random effects in the model by comparing the model to a model fitted with just the fixed effects and excluding the random effects.  (they are the same in depth)
```{r}

model.fixed = gls(CVH~gender+insurance+age+education,
                  data=data,
                  method="ML")

anova(model.fixed,mixed)
```
We can see that the random effects are significant, and the mixed model has smaller AIC and BIC and larger loglik


```{r cars}
summary(mixed)
```

To conclude, factors related to a favorable CVH score included insurance covered, younger age, female sex, and a higher level of education.

So the answer to the question we brought up is yes, women tend to have a better cardiovascular health condition than men in the US. 

